---
Codebook : Getting and Cleaning data project
output: codebook.md
---

## Preparation 

```{r }
#############    PREPARARTION     ###############
# 1. set work directory
# 2. include all necessary packages
# 3. download all files and dataset
# 4. Unzip data files

setwd("/Users/xinsui/Desktop/Rtraining/GandCData")
path <- getwd()
library(data.table)
library(reshape2)


fileURL <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
destfile <- "dataset.zip"
download.file(fileURL, file.path(path, destfile))
unzip(file.path(path,destfile), files = NULL, list = FALSE, overwrite = TRUE,junkpaths = FALSE, exdir = ".", unzip = "internal",
      setTimes = FALSE )
datapath = file.path(path,"/UCI HAR Dataset")
list.files(datapath, recursive = TRUE)

```

## Reading Data from raw files

### 1.Activity labels

```{r }
# Activity Labels 
activityLabels <- fread(file.path(datapath, "activity_labels.txt"))
setnames (activityLabels, c("V1", "V2"), c("act_num", "Activity"))
activityLabels
```

### 2.Feature names list

```{r }
features <- fread(file.path(datapath, "features.txt"))
setnames(features, c("V1", "V2"), c("feature_num", "feature_name"))

# Get only the data on mean and std. dev features
features <- features[grepl("mean\\(\\)|std\\(\\)", feature_name)]
features$feature_num <- features[, paste0("V", feature_num)]
wanted <- features$feature_num
head(features)
```

### 3.Read all training and test files
```{r }
# Read Train and Test Data
dtSubTrain <- fread(file.path(datapath, "train", "Subject_train.txt"))
dtSubTest <- fread(file.path(datapath, "test", "Subject_test.txt"))

dtTrain <- fread(file.path(datapath, "train", "X_train.txt"), select = wanted)
dtTest <- fread(file.path(datapath, "test", "X_test.txt"), select = wanted)
dtActTrain <- fread(file.path(datapath, "train", "Y_train.txt"))
dtActTest <- fread(file.path(datapath, "test", "Y_test.txt"))  

# Make descriptive variable names 
setnames(dtSubTrain, "V1", "subject_id")
setnames(dtActTrain, "V1", "act_num")
setnames(dtSubTest, "V1", "subject_id")
setnames(dtActTest, "V1", "act_num")

```

### 4.Merge training and test data 
```{r}
dtTrain <- cbind(dtTrain, dtSubTrain, dtActTrain)
dtTest <- cbind(dtTest, dtSubTest, dtActTest)
dtData = rbind(dtTrain, dtTest)
setkey(dtData, subject_id, act_num)

head(dtData)
```

### 5.descriptive activity names to name the activities in the data set
```{r}
dtData <- merge(dtData, activityLabels, by = "act_num", all.x = TRUE)
setkey(dtData, subject_id, act_num, Activity)
```
### Appropriately labels the data set with descriptive variable names.
```{r}
# Get descriptive feature names.
dtMelted <- data.table(melt(dtData, key(dtData), variable.name = "feature_num"))
dtMelted <- merge(dtMelted, features[, list(feature_num, feature_name)], by = "feature_num", 
            all.x = TRUE)

# Separate different feature variables 

## Features with 1 category
dtMelted$Jerk <- factor(grepl("Jerk",dtMelted$feature_name), labels = c(NA, "Jerk"))
dtMelted$Magnitude <- factor(grepl("Mag",dtMelted$feature_name), labels = c(NA, "Magnitude"))
## Features with 2 categories
n <- 2
y <- matrix(seq(1, n), nrow = n)
x <- matrix(c(grepl("^t", dtMelted$feature_name), grepl("^f",dtMelted$feature_name)), ncol = nrow(y))
dtMelted$Domain <- factor(x %*% y, labels = c("time", "frequency"))
x <- matrix(c(grepl("Acc", dtMelted$feature_name), grepl("Gyro",dtMelted$feature_name)), ncol = nrow(y))
dtMelted$Signals <- factor(x %*% y, labels = c("Accelerometer", "Gyroscope"))
x <- matrix(c(grepl("BodyAcc", dtMelted$feature_name), grepl("GravityAcc", dtMelted$feature_name)), ncol = nrow(y))
dtMelted$Motion <- factor(x %*% y, labels = c(NA, "Body", "Gravity"))
x <- matrix(c(grepl("mean()",dtMelted$feature_name), grepl("std()",dtMelted$feature_name)), ncol = nrow(y))
dtMelted$Variable <- factor(x %*% y, labels = c("Mean", "SD"))

## Features with 3 categories
n <- 3
y <- matrix(seq(1, n), nrow = n)
x <- matrix(c(grepl("-X",dtMelted$feature_name), grepl("-Y",dtMelted$feature_name), grepl("-Z",dtMelted$feature_name)), ncol = nrow(y))
dtMelted$Axis <- factor(x %*% y, labels = c(NA, "X", "Y", "Z"))

head(dtMelted)

```
### create tidy data set with the average of each variable for each activity and each subject.
```{r}

setkey(dtMelted, subject_id, Activity, Domain, Signals, Motion, 
       Jerk, Magnitude, Variable, Axis)
dtTidy <- dtMelted[, list( average = mean(value)), by = key(dtMelted)]


```


### write into a text file
```{r}
write.table(dtTidy, "tidy.txt", row.names = FALSE, quote = FALSE)
```
